---
title: "609project.rmd"
output:
  html_document: default
  pdf_document: default
---

##  Introduction
#### Hypothesis:
H1: The severity of generalized anxiety is related to the increased impacts of COVID-19 on ability to meet financial obligations or essential needs(A). 

H2: The severity of generalized anxiety is related to the increased person’s time spending on the internet(I). 
H3: The severity of generalized anxiety is related to the increased household food insecurity (H). 

H4: The severity of generalized anxiety is related to the increased effect of A, H, and I. 
\

## Data Overview
Dependent Variable: ANXDVGAD (Generalized Anxiety Severity Score)

Independent Variable: 

1.	Key Variables:
a.	LM_40 (COVID-19 impacts ability meet financial obligations or essential needs)
b.	PFSCDV (Household food insecurity)
c.	BH_40F (Spending time on the internet)
\

2.	Control Variables:
a.	AGEGRP (Age group of respondent)
b.	SEX (Sex of respondent)
c.	PCHILD (Child under 18 on May 4th, 2020 resides in dwelling)
\

3. Reorder some variables as follows:
a. LM_40 (1: No impact; 2: Minor impact; 3: Moderate impact; 4: Major impact)
b. BH_40F (1: Decrease; 2: No change; 3: Increase)
c. PFSCDV (1: Food secure; 2: Marginally food insecure; 3: Moderately food insecure; 4: Severely food insecure)
d. Sex (0: Male; 1: Female)
\
\

## Loading objects and data
```{r cache=T}
packages=c("caret","ggplot2","tidyverse","dplyr","corrplot","DescTools","zoo","lmtest","MASS","stargazer","erer","car","brant")
# Now load or install & load all
package.check <- lapply(packages,FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
}
)

setwd("/Users/zoe/Desktop")
mh=read.csv("mh.csv", header=T)
mh=as_tibble(mh)

```

## Data Visualization and Exploration
````{r}
# changing the names of column
mh_rename = setNames(mh, c("mh_index", "financ_stress", "internet", "food_insec", "age", "children", "sex"))
summary(mh_rename)

ggplot(data=mh_rename) + geom_bar(mapping = aes(x = financ_stress),color = 'skyblue4',fill='skyblue4')
ggplot(data=mh_rename) + geom_bar(mapping = aes(x = internet),color = 'skyblue4',fill='skyblue4')
ggplot(data=mh_rename) + geom_bar(mapping = aes(x = food_insec),color = 'skyblue4',fill='skyblue4')
ggplot(data=mh_rename) + geom_bar(mapping = aes(x = age),color = 'skyblue4',fill='skyblue4')
ggplot(data=mh_rename) + geom_bar(mapping = aes(x = children),color = 'skyblue4',fill='skyblue4')
ggplot(data=mh_rename) + geom_bar(mapping = aes(x = sex),color = 'skyblue4',fill='skyblue4')
ggplot(data=mh_rename,aes(x = mh_index)) + geom_bar(color = 'skyblue4',fill='skyblue4')

# Scatter Plot
plot(mh_rename)

# Skewness and Kurtosis
install.packages("moments")
library(moments)
skewness(mh_rename$mh_index)


# Correlation Plot
mh_corr = cor(mh_rename,method = 'spearman')
corrplot(mh_corr, # correlation matrix
         order = "hclust") 
corrplot(mh_corr, order = "hclust") 

```

## Data visualization after grouping 
```{r}
# Data grouping
mh_convert = mh_rename %>%
  mutate(mh_index_convert = case_when(mh_index <= 4 ~ "1",
                                      mh_index == 5 ~ "2",
                                      mh_index == 6 ~ "2",
                                      mh_index == 7 ~ "2",
                                      mh_index == 8 ~ "2",
                                      mh_index == 9 ~ "2",
                                      mh_index == 10 ~ "3",
                                      mh_index == 11 ~ "3",
                                      mh_index == 12 ~ "3",
                                      mh_index == 13 ~ "3",
                                      mh_index == 14 ~ "3",
                                      mh_index >= 15 ~ "4",))

# Financial obligation
mh1=mh_convert %>%
  select(financ_stress,mh_index_convert) %>%
  group_by(financ_stress) %>%
  count(mh_index_convert) %>%
  mutate(percentage=n/sum(n))
ggplot(mh1, aes(fill=factor(mh_index_convert), x=financ_stress, y=percentage))+geom_bar(stat="identity")

# internet
mh2=mh_convert %>%
  select(internet,mh_index_convert) %>%
  group_by(internet) %>%
  count(mh_index_convert) %>%
  mutate(percentage=n/sum(n))
ggplot(mh2, aes(fill=factor(mh_index_convert), x=internet, y=percentage))+geom_bar(stat="identity")

# food insecurity
mh3=mh_convert %>%
  select(food_insec,mh_index_convert) %>%
  group_by(food_insec) %>%
  count(mh_index_convert) %>%
  mutate(percentage=n/sum(n))
ggplot(mh3, aes(fill=factor(mh_index_convert), x=food_insec, y=percentage))+geom_bar(stat="identity")

# age group
mh4=mh_convert %>%
  select(age,mh_index_convert) %>%
  group_by(age) %>%
  count(mh_index_convert) %>%
  mutate(percentage=n/sum(n))
ggplot(mh4, aes(fill=factor(mh_index_convert), x=age, y=percentage))+geom_bar(stat="identity")

# child under 18
mh5=mh_convert %>%
  select(children,mh_index_convert) %>%
  group_by(children) %>%
  count(mh_index_convert) %>%
  mutate(percentage=n/sum(n))
ggplot(mh5, aes(fill=factor(mh_index_convert), x=children, y=percentage))+geom_bar(stat="identity")

# sex
mh6=mh_convert %>%
  select(sex,mh_index_convert) %>%
  group_by(sex) %>%
  count(mh_index_convert) %>%
  mutate(percentage=n/sum(n))
ggplot(mh6, aes(fill=factor(mh_index_convert), x=sex, y=percentage))+geom_bar(stat="identity")

```

## Multiple regression model
```{r}
# Method 1. Original dataset

mh.lm1 = lm(mh_index ~ financ_stress + age + children + sex, data = mh_rename)
summary(mh.lm1)

mh.lm2 = lm(mh_index ~ internet + food_insec + age + children + sex, data = mh_rename)
summary(mh.lm2)

mh.lm3 = lm(mh_index ~ food_insec + age + children + sex, data = mh_rename)
summary(mh.lm3)

mh.lm4 = lm(mh_index ~ financ_stress + internet + food_insec + age + children + sex, data = mh_rename)
summary(mh.lm4)


# Method 2. log transformation on my DV. 
# mh_index can't be 0
# increment column of mh_index by 1. 
mh_add = mh_rename %>% mutate(mh_index = mh_index + 1)
# transform mh_index to log
mh_trans.lm = lm(log(mh_index) ~ financ_stress + internet + food_insec + age + children + sex, data = mh_add)
summary(mh_trans.lm)

```

## Normality test
```{r}
par(mfrow = c(2, 2))
plot(mh.lm4)

par(mfrow = c(2, 2))
plot(mh_trans.lm)

hist(residuals(mh.lm4))
boxplot(residuals(mh.lm4))
shapiro.test(residuals(mh.lm4))
             
hist(residuals(mh_trans.lm))
boxplot(residuals(mh_trans.lm))
shapiro.test(residuals(mh_trans.lm))

```

## Residual check and heteroskedasticity
```{r}
# The Breusch-Pagan Test for heteroskedasticity
# H0: There is no heteroskedasticity
# H1: There is heteroskedasticity

bptest(mh.lm4)
bptest(mh_trans.lm)
# We can reject the null hypothesis, so in both models, we do have heteroskedasticity issue. 

# Resolving heteroskedasticity
# For untransformed data
mh_resolve = lm(mh_index ~ financ_stress + internet + food_insec + age + children + sex, data = mh_rename)
mh$resi = mh_resolve$residuals
mh_varfunc = lm(log(mh$resi^2) ~ log(financ_stress + internet + food_insec + age + children + sex), data = mh_rename)
mh_vari <- exp(mh_varfunc$fitted.values)
mh.gls <- lm(mh_index ~ financ_stress + internet + food_insec + age + children + sex, weights = 1/sqrt(mh_vari), data = mh_rename)
summary(mh.gls)

bptest(mh.gls)

# Still have heteroskedasticity issue. 

# For log transformed data
mh_resolve = lm(log(mh_index) ~ financ_stress + internet + food_insec + age + children + sex, data = mh_add)
mh$resi = mh_resolve$residuals
mh_varfunc = lm(log(mh$resi^2) ~ log(financ_stress + internet + food_insec + age + children + sex), data = mh_add )
mh_vari <- exp(mh_varfunc$fitted.values)
mh.gls <- lm(log(mh_index) ~ financ_stress + internet + food_insec + age + children + sex, weights = 1/sqrt(mh_vari), data = mh_add )
summary(mh.gls)

bptest(mh.gls)

# Still have heteroskedasticity issue. 

```

## Ordered Logit Model
```{r}
# Method 3: Ordered Logit Model

# Grouping Reference: "INSTRUCTION MANUAL Instructions for Patient Health Questionnaire (PHQ) and GAD-7 Measures"
# 0-4 minimal anxiety
# 5-9 mild anxiety
# 10-14 moderate anxiety
# >=15 severe anxiety

mh_logit_all = polr(as.factor(mh_index_convert) ~ financ_stress + internet + food_insec + age + children + sex, data = mh_convert, Hess=TRUE)

PseudoR2(mh_logit_all, which = NULL)

# summary(mh_logit.lm)
# 
# summary_table <- coef(summary(mh_logit.lm))
# pval <- pnorm(abs(summary_table[, "t value"]),lower.tail = FALSE)* 2
# summary_table <- cbind(summary_table, "p value" = round(pval,3))
# summary_table


# Getting the coefficient and p-value
mh_logit_all.coef = data.frame(coef(summary(mh_logit_all)))
mh_logit_all.coef$pval = round((pnorm(abs(mh_logit_all.coef$t.value), lower.tail = FALSE) * 2),2)
mh_logit_all.coef

library(stargazer)
stargazer(mh_logit_all, type="text")

# Odds ratios
mh_logit_all.or=exp(coef(mh_logit_all))
mh_logit_all.or

stargazer(mh_logit_all, type="text", coef=list(mh_logit_all.or), p.auto=FALSE)

# Interpretation: 
# For financial obligation, keep all variables constant, when financ_stress increases by 1 unit, it is 1.583 times more likely to be in a higher category of mental health index (more severe mental health issue). In other words, the odds of moving to a higher category in the outcome variable is 58.3% when financ_stress move 1 unit (1.583 - 1). The coefficient is significant. 

# For children, the coefficient is not significant. Here, the children is the control variable. 

```

## Ordered Probit Model
```{r}
# Method 4: Ordered Probit Model
# install.packages("aod")
# library(aod)
# 
# mh_logit.lm = glm(as.factor(mh_index) ~ financ_stress + internet + food_insec + age + children + sex, family = "binomial",data = mh_rename)
# 
# summary(mh_logit.lm)
# 
# PseudoR2(mh_logit.lm, which = NULL)

mh_probit_all = polr(as.factor(mh_index_convert) ~ financ_stress + internet + food_insec + age + children + sex, data = mh_convert, Hess=TRUE, method = "probit")

PseudoR2(mh_probit_all, which = NULL)

mh_probit_all.coef = data.frame(coef(summary(mh_probit_all)))
mh_probit_all.coef$pval = round((pnorm(abs(mh_probit_all.coef$t.value), lower.tail = FALSE) * 2),2)
mh_probit_all.coef

library(stargazer)
stargazer(mh_probit_all, type= "text")

# Interpretation:
# For financial obligation, keep all variables constant, when financ_stress increases by 1 unit, the z-score of mh_index increases by 0.269. The coefficient is significant. 

# For children, the coefficient is not significant. 
```

## Split data and build models
```{r}
# Split data into training set and test set
samplesize = 0.60*nrow(mh_convert)
set.seed(100)
index = sample(seq_len(nrow(mh_convert)), size = samplesize)

#Creating training and test set 
mh_train = mh_convert[index,]
mh_test = mh_convert[-index,]

# Build model using train set

# Logit1

logit1 = polr(as.factor(mh_index_convert) ~ financ_stress + age + children + sex, data = mh_train, Hess=TRUE)
summary(logit1)
table1 = coef(summary(logit1))
p = pnorm(abs(table1[,"t value"]), lower.tail = FALSE) * 2
(ctable1 = cbind(table1, "p value" = p))

stargazer(logit4, type="text")

# Odds ratios

ci = confint(logit1)
confint.default(logit1)
exp(coef(logit1))
exp(cbind(OR=coef(logit1), ci))


# Logit2

logit2 = polr(as.factor(mh_index_convert) ~ internet + age + children + sex, data = mh_train, Hess=TRUE)
summary(logit2)
table2 = coef(summary(logit2))
p = pnorm(abs(table2[,"t value"]), lower.tail = FALSE) * 2
(ctable1 = cbind(table2, "p value" = p))

library(stargazer)
stargazer(logit2, type="text")

# Odds ratios
logit2.or=exp(coef(logit2))
logit2.or

stargazer(logit2, type="text", coef=list(logit2.or), p.auto=FALSE)


# Logit3

logit3 = polr(as.factor(mh_index_convert) ~ food_insec + age + children + sex, data = mh_train, Hess=TRUE)
summary(logit3)

table3 = coef(summary(logit3))
p = pnorm(abs(table3[,"t value"]), lower.tail = FALSE) * 2
(ctable1 = cbind(table3, "p value" = p))

library(stargazer)
stargazer(logit3, type="text")

# Odds ratios
logit3.or=exp(coef(logit3))
logit3.or

stargazer(logit3, type="text", coef=list(logit3.or), p.auto=FALSE)

# Logit4

logit4 = polr(as.factor(mh_index_convert) ~ financ_stress + internet + food_insec + age + children + sex, data = mh_train, Hess=TRUE)
summary(logit4)

table4 = coef(summary(logit4))
p = pnorm(abs(table4[,"t value"]), lower.tail = FALSE) * 2
(ctable4 = cbind(table4, "p value" = p))

stargazer(logit4, type="text")

# Odds ratios
logit4.or=exp(coef(logit4))
logit4.or

stargazer(logit4, type="text", coef=list(logit4.or), p.auto=FALSE)

# Interpretation:
# According to AIC, logit 4 is the best model of the four with smallest AIC. 

# The interpretation for the coefficients is as follows. For example, holding everything else constant, an increase in value of financial obligation by one unit increase the expected value of mh_index in log odds by 0.4635. 

# 1|2 probability intercept takes value of 2.1843, indicating that the expected odds of identifying in 1 category, when other variables assume a value of zero, is 2.1843. 

(prob_1 = exp(2.1843)/(1+exp(2.1843)))
(prob_2 = exp(3.963)/(1+exp(3.863)))


# The expected probability of identifying 1 category of mh_index, when other variables assume a value of zero, is 0.8988.


# Compute confusion table and misclassification error

# logit 1
predict_index1 = predict(logit1, mh_test)
table(mh_test$mh_index_convert, predict_index1)
mean(as.character(mh_test$mh_index_convert) != as.character(predict_index1))


# logit 2
predict_index2 = predict(logit2, mh_test)
table(mh_test$mh_index_convert, predict_index2)
mean(as.character(mh_test$mh_index_convert) != as.character(predict_index2))

# logit 3
predict_index3 = predict(logit3, mh_test)
table(mh_test$mh_index_convert, predict_index3)
mean(as.character(mh_test$mh_index_convert) != as.character(predict_index3))

# logit 4
predict_index4 = predict(logit4, mh_test)
table(mh_test$mh_index_convert, predict_index4)
mean(as.character(mh_test$mh_index_convert) != as.character(predict_index4))

# Interpretation:
# In the test dataset, 847 times index = 1 is identified correctly. 
# Using the confusion matrix, we find that the misclassification error for our models are all around 40%.

```

## Probability for each category of DV.
```{r}
logit4.coef = data.frame(coef(summary(logit4)))
logit4.coef$pval = round((pnorm(abs(logit4.coef$t.value), lower.tail = FALSE) * 2),5)
logit4.coef

dv_4 = mh_train$mh_index[mh_train$mh_index_convert == 4]
odds = (length(dv_4)/length(mh_train$mh_index_convert))
beta0 = log(odds*(1-odds))

logit4.or = exp(coef(logit4))
stargazer(logit4, type="text", coef=list(logit4.or), p.auto=FALSE, out="text")
logit4.pred = predict(logit4, type="probs")
summary(logit4.pred)

setup = data.frame(sex=mh_train$sex,
                    age=mh_train$age,
                    children=mh_train$children,
                    financ_stress=mh_train$financ_stress,
                    internet=mh_train$internet,
                    food_insec=mh_train$food_insec)

setup[, c("pred.prob")] = predict(logit4, newdata = setup, type = "probs")
setup[1:50,]

```
## Marginal Effect
```{r}
library(margins)
library(broom)
margin1 = glm((as.numeric(mh_index_convert) == 1 ) ~ financ_stress + internet + food_insec + age + children + sex, family = "binomial"(link="logit"), data = mh_train)
margin2 = glm((as.numeric(mh_index_convert) == 2) ~ financ_stress + internet + food_insec + age + children + sex, family = "binomial"(link="logit"), data = mh_train)
margin3 = glm((as.numeric(mh_index_convert) == 3) ~ financ_stress + internet + food_insec + age + children + sex, family = "binomial"(link="logit"), data = mh_train)
margin4 = glm((as.numeric(mh_index_convert) == 4) ~ financ_stress + internet + food_insec + age + children + sex, family = "binomial"(link="logit"), data = mh_train)

logitmargins1 = margins(margin1)
logitmargins2 = margins(margin2)
logitmargins3 = margins(margin3)
logitmargins4 = margins(margin4)

```

## Checking ordered logit assumption
```{r}
# 1. Multi-collinearity
h4.model = lm(scale(as.numeric(mh_index_convert)) ~ financ_stress + internet + food_insec + age + children + sex, data = mh_train)
vif(h4.model)
# Pass

# 2. Proportional odds
brant(logit4)
# Pass

```

